Calendar-based graphics for visualizing people's daily schedules

### 2017-09-06 main.Rmd
minor tweaks for resubmission

---
author:
  - name: Earo Wang
    affiliation: Monash University
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{earo.wang@monash.edu}
  - name: Dianne Cook
    affiliation: Monash University
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{dicook@monash.edu}
  - name: Rob J Hyndman
    affiliation: Monash University
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{rob.hyndman@monash.edu}
title:
  formatted: "Calendar-based graphics for visualizing people's daily schedules"
  plain: "Calendar-based graphics for visualizing people's daily schedules"
abstract: >
  This paper describes a \code{frame\_calendar} function that organizes and displays temporal data, collected on sub-daily resolution, into a calendar layout. Calendars are broadly used in society to display temporal information, and events. The \code{frame\_calendar} uses linear algebra on the date variable to create the layout. It utilizes the grammar of graphics to create the plots inside each cell, and thus synchronizes neatly with \pkg{ggplot2} graphics. The motivating application is studying pedestrian behavior in Melbourne, Australia, based on counts which are captured at hourly intervals by sensors scattered around the city. Faceting by the usual features such as day and month, was insufficient to examine the behavior. Making displays on a monthly calendar format helps to understand pedestrian patterns relative to events such as work days, weekends, holidays, and special events. The layout algorithm has several format options and variations. It is implemented in the \proglang{R} package \pkg{sugrrants}.

keywords:
  # at least one keyword must be supplied
  formatted: [data visualization, statistical graphics, time series, grammar of graphics, "\\proglang{R}"]
  plain:     [data visualization, statistical graphics, time series, grammar of graphics, R]
preamble: >
  \usepackage{amsmath}
output: 
  bookdown::pdf_book:
    base_format: rticles::jss_article
  citation_package: biblatex
bibliography: ["references.bib", "packages.bib"]
---

@@

We develop a method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout. The calendar format is created using linear algebra, and fits into the \pkg{tidyverse} [@R-tidyverse] and grammar of graphics [@wickham2009ggplot2] framework for exploring data. The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events. This work was originally motivated by studying foot traffic in the city of Melbourne, Australia [@ped]. There have been 43 sensors installed that count pedestrians every hour across the downtown area (Figure \ref{fig:ped-map}). The dataset can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management. A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns. The faceted plots in Figure \ref{fig:time-series-plot}, give an overall picture of the foot traffic at 3 different sensors over 2016. Further faceting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse of the daily and sub-daily pedestrian patterns. The sensor data, like many temporal datasets on human behaviors, lends itself to a number of exploratory data visualization challenges:

>>>>

---
author:
  - name: Earo Wang
    affiliation: Monash University
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{earo.wang@monash.edu}
  - name: Dianne Cook
    affiliation: Monash University
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{dicook@monash.edu}
  - name: Rob J Hyndman
    affiliation: Monash University
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{rob.hyndman@monash.edu}
title:
  formatted: "Calendar-based graphics for visualizing people's daily schedules"
  plain: "Calendar-based graphics for visualizing people's daily schedules"
abstract: >
  This paper describes a \code{frame\_calendar} function that organizes and displays temporal data, collected on sub-daily resolution, into a calendar layout. Calendars are broadly used in society to display temporal information, and events. The \code{frame\_calendar} uses linear algebra on the date variable to create the layout. It utilizes the grammar of graphics to create the plots inside each cell, and thus synchronizes neatly with \pkg{ggplot2} graphics. The motivating application is studying pedestrian behavior in Melbourne, Australia, based on counts which are captured at hourly intervals by sensors scattered around the city. Faceting by the usual features such as day and month, was insufficient to examine the behavior. Making displays on a monthly calendar format helps to understand pedestrian patterns relative to events such as work days, weekends, holidays, and special events. The layout algorithm has several format options and variations. It is implemented in the \proglang{R} package \pkg{sugrrants}.
keywords:
  # at least one keyword must be supplied
  formatted: [data visualization, statistical graphics, time series, grammar of graphics, "\\proglang{R}"]
  plain:     [data visualization, statistical graphics, time series, grammar of graphics, R]
preamble: >
  \usepackage{amsmath}
output: 
  bookdown::pdf_book:
    base_format: rticles::jss_article
  citation_package: biblatex
bibliography: ["references.bib", "packages.bib"]
---

@@

We develop a method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout. The calendar format is created using linear algebra, fitting nicely into the \pkg{tidyverse} [@R-tidyverse] suite. In particular, the resulting coordinates synchronize with \pkg{ggplot2} [@R-ggplot2] graphics for exploring data. The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events. This work was originally motivated by studying foot traffic in the city of Melbourne, Australia [@ped]. There have been 43 sensors installed that count pedestrians every hour across the downtown area (Figure \ref{fig:ped-map}). The dataset can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management. A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns. The faceted plots in Figure \ref{fig:time-series-plot}, give an overall picture of the foot traffic at 3 different sensors over 2016. Further faceting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse of the daily and sub-daily pedestrian patterns. The sensor data, like many temporal datasets on human behaviors, lends itself to a number of exploratory data visualization challenges:

### 2017-09-06 main.tex
minor tweaks for resubmission

We develop a method for organizing and visualizing temporal data,
collected at sub-daily intervals, into a calendar layout. The calendar
format is created using linear algebra, and fits into the
\pkg{tidyverse} \citep{R-tidyverse} and grammar of graphics
\citep{wickham2009ggplot2} framework for exploring data. The purpose of
the calendar-based visualization is to provide insights into people's
daily schedules relative to events such as work days, weekends,
holidays, and special events. This work was originally motivated by
studying foot traffic in the city of Melbourne, Australia \citep{ped}.
There have been 43 sensors installed that count pedestrians every hour
across the downtown area (Figure \ref{fig:ped-map}). The dataset can
shed light on people's daily rhythms, and assist the city administration
and local businesses with event planning and operational management. A
routine examination of the data would involve constructing conventional
time series plots to catch a glimpse of temporal patterns. The faceted
plots in Figure \ref{fig:time-series-plot}, give an overall picture of
the foot traffic at 3 different sensors over 2016. Further faceting by
day of the week (Figure \ref{fig:facet-time}) provides a better glimpse
of the daily and sub-daily pedestrian patterns. The sensor data, like
many temporal datasets on human behaviors, lends itself to a number of
exploratory data visualization challenges:

>>>>

We develop a method for organizing and visualizing temporal data,
collected at sub-daily intervals, into a calendar layout. The calendar
format is created using linear algebra, fitting nicely into the
\pkg{tidyverse} \citep{R-tidyverse} suite. In particular, the resulting
coordinates synchronize with \pkg{ggplot2} \citep{R-ggplot2} graphics
for exploring data. The purpose of the calendar-based visualization is
to provide insights into people's daily schedules relative to events
such as work days, weekends, holidays, and special events. This work was
originally motivated by studying foot traffic in the city of Melbourne,
Australia \citep{ped}. There have been 43 sensors installed that count
pedestrians every hour across the downtown area (Figure
\ref{fig:ped-map}). The dataset can shed light on people's daily
rhythms, and assist the city administration and local businesses with
event planning and operational management. A routine examination of the
data would involve constructing conventional time series plots to catch
a glimpse of temporal patterns. The faceted plots in Figure
\ref{fig:time-series-plot}, give an overall picture of the foot traffic
at 3 different sensors over 2016. Further faceting by day of the week
(Figure \ref{fig:facet-time}) provides a better glimpse of the daily and
sub-daily pedestrian patterns. The sensor data, like many temporal
datasets on human behaviors, lends itself to a number of exploratory
data visualization challenges:
